Red Wine Quality by Robert Bornemann
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(gridExtra)
library(GGally)
library(corrplot)
library(psych)
library(memisc)
library(dplyr)
library(car)

```

```{r echo=FALSE, Load_the_Data}
# Load the Data

dataset = read.csv("wineQualityReds.csv")
options(warn=-1)

```

This is a R Markdown EDA report on a dataset containing 1599 red wine samples. 
The goal was to clean and understand the dataset and explore what chemical 
predicators define wine quality. The dataset used in this report is 
publicly available for research through:

[@Elsevier] http://dx.doi.org/10.1016/j.dss.2009.05.016 \
[Pre-press (pdf)] http://www3.dsi.uminho.pt/pcortez/winequality09.pdf \
[bib] http://www3.dsi.uminho.pt/pcortez/dss09.bib

# Univariate Plots Section

Our dataset consists of 13 - 1 variables, with 1599 observations. I excluded the
variable X in the analysis as it is simply another ID column that R Studio 
doesn't need.

```{r echo=FALSE, Univariate_Plots}
dim(dataset)
str(dataset)
summary(dataset)
```

Number of missing values in the dataset:

```{r echo=FALSE}
sum(is.na(dataset))

dataset$quality.factor = factor(dataset$quality)

ggplot(data = dataset, aes(quality)) +
  geom_histogram(binwidth = .5, col="black", fill="grey", alpha = .8) +
  scale_x_continuous(breaks=seq(0,8,1))

```

I decided to explore the distribution of quality ratings first. The histogram 
reveals that most red wines in the dataset are of medium quality (μ 5.63) on a 
scale between 0 (very bad) and 10 (very excellent). The worse red wine has a 
quality of 3 whereas the best red wine in the dataset has a max. quality score 
of 8. The distribution is slightly skewed to the left suggesting that there 
should be a higher number of wines from medium to higher quality in the dataset. 

The question we want to ask is which of the follwing chemical properties can 
help us explain the quality ratings of red wines based on sensory data by wine 
experts. Lets try to find out!

```{r echo=FALSE}
# plotting all acids together first without limits and transformations

plot1 = ggplot(data = dataset, aes(fixed.acidity)) +
  geom_histogram(binwidth = .5, col="black", fill="grey", alpha = .8) +
  scale_x_continuous(breaks=seq(4,16,1))

plot2 = ggplot(data = dataset, aes(volatile.acidity)) +
  geom_histogram(binwidth = .05, col="black", fill="grey", alpha = .8) +
  scale_x_continuous(breaks=seq(0,1.6,.1))

plot3 = ggplot(data = dataset, aes(citric.acid)) +
  geom_histogram(binwidth = .005, col="black", fill="grey", alpha = .8)

grid.arrange(plot1, plot2, plot3)

```

All the chemical properties describing different acids in the 3 histrograms 
above seem to be right skewed in various forms. A closer look however reveals 
that the distrutions of fixed and volatile acidity are most likely skewed 
because of some outliers. My idea was to trim both of the variables distrutions 
in order to make them approximately more normal in distribution for further 
analysis. 

Moreover, I am attempting to transform the citric acid variable using the 
squareroot method in order to get a clearer view of its distrution. The 
distrutions looks like a waveform peaking and declining at various points of the
scale. Zooming in might help to understand this pattern a little better.

```{r echo=FALSE}
         
ggplot(data = dataset, aes(fixed.acidity)) +
  geom_histogram(binwidth = .5, col="black", fill="grey", alpha = .8) +
  scale_x_continuous(breaks=seq(4,16,1), limits = c(4, quantile(dataset$fixed.acidity, .99)))

summary(dataset$fixed.acidity)

ggplot(data = dataset, aes(volatile.acidity)) +
  geom_histogram(binwidth = .05, col="black", fill="grey", alpha = .8) +
  scale_x_continuous(breaks=seq(0,1.6,.1), limits = c(0, 1))

summary(dataset$volatile.acidity)

ggplot(data = dataset, aes(citric.acid)) +
  geom_histogram(binwidth = .005, col="black", fill="grey", alpha = .8) +
  scale_x_sqrt()

ggplot(data = dataset, aes(citric.acid)) +
  geom_histogram(binwidth = .005, col="black", fill="grey", alpha = .8) +
  scale_x_sqrt(limits = c(.01, 1))

summary(dataset$citric.acid)
  
```

A closer look at the histograms above illustrates how the distrutions of fixed 
and volatile acidity became more normal in distribution through limiting the 
x-axis. 

Citric acid seems to be different. We can see that a significant part of red 
wines has a very little or no amount of citric acid. About 8% of red wines 
contain no citric acid, which makes adding 'freshness' and flavor to wines 
using citric acid rather look optional [Cortez et al., 2009]. On the other hand 
we can also confirm the bimodal wavepattern from the first histogram and see 
that the counts are peaking around 0.25 and 0.50. Perhaps adding citric acid is 
a rather delicate process and mastering this techniques adds to the overall 
quality score? Hopefully a correlation analysis will clarify these observations 
further.

```{r echo=FALSE}

plot4 = ggplot(data = dataset, aes(residual.sugar)) +
  geom_histogram(binwidth = .5, col="black", fill="grey", alpha = .8) +
  scale_x_continuous()

plot5 = ggplot(data = dataset, aes(chlorides)) +
  geom_histogram(binwidth = .025, col="black", fill="grey", alpha = .8) +
  scale_x_continuous()

grid.arrange(plot4, plot5)
```

Looking at sugars and salts in wine we can also observe a right skewed 
distribution. Moreover, outliers are stretching both distributions to more than 
double the size of where the majoroity of the data sits. I am attempting to 
transform both variables using a log10 method in order to get a clearer view of 
the distrutions as well as limiting the histogram by ignoring the 1% outliers on 
the very right of the scale. 

```{r echo=FALSE}
ggplot(data = dataset, aes(residual.sugar)) +
  geom_histogram(binwidth = .05, col="black", fill="grey", alpha = .8) +
  scale_x_log10(limits = quantile(dataset$residual.sugar, c(0, .99)))

summary(dataset$residual.sugar)

ggplot(data = dataset, aes(chlorides)) +
  geom_histogram(binwidth = .1, col="black", fill="grey", alpha = .8) +
  scale_x_log10(limits = quantile(dataset$chlorides, c(0, .99)))

summary(dataset$chlorides)
```

The transformatins applied helped to understand both distrutions better. While
sugar still seems to be slightly skewed to the right we can observe that 
clorides peak very sharp at a very particular point on the scale. It seems that
there is far more variation in the amount of sugar and less so in salts. I am, 
at this point not yet very sure about my intuition on these variables. However,
I feel there might be a relationship between sugar and salt ratios and their 
effect on the quality rating. 

```{r echo=FALSE}

plot6 = ggplot(data = dataset, aes(free.sulfur.dioxide)) +
  geom_histogram(binwidth = 2.5, col="black", fill="grey", alpha = .8) +
  scale_x_continuous()

plot7 = ggplot(data = dataset, aes(total.sulfur.dioxide)) +
  geom_histogram(binwidth = 10, col="black", fill="grey", alpha = .8) +
  scale_x_continuous()

grid.arrange(plot6, plot7)

```

As for many of the other chemical properties in the dataset free.sulfur.dioxide
and total.sulfur.dioxide are right skewed. The existence of these properties in
wine is important as it prevents microbial growth and the oxidation. However, 
concentrations over 50 ppm become evident in the nose and taste and are 
therefore most likely less disirable. Perhaps this can explain some of the 
variance in the lower quality ranks. Therefore, my intution at this point is 
that the distributions are right skewed because of this. 

I am also attempting to transform both variables using a log10 method in order 
to get a clearer view of the distrutions. 

```{r echo=FALSE}

ggplot(data = dataset, aes(free.sulfur.dioxide)) +
  geom_histogram(binwidth = .05, col="black", fill="grey", alpha = .8) +
  scale_x_log10()

summary(dataset$free.sulfur.dioxide)

ggplot(data = dataset, aes(total.sulfur.dioxide)) +
  geom_histogram(binwidth = .05, col="black", fill="grey", alpha = .8) +
  scale_x_log10()

summary(dataset$total.sulfur.dioxide)
```

Interestingly, the log10 method helped significantly to transform the variables 
to a more normal distribution. While total.sulfur.dioxide shows a nearly 
bell-type pattern in the histogram we can observe that free.sulfur.dioxide is 
distributed in a more scattered way. There are some outliers that contain pretty 
much no or very little free.sulfur.dioxide. In further evalutation of the same 
variable we see that its presence throughout the distribution is rather 
irregular. However, due to the nature of the variable itself I have a very 
limited intuition about its significance in relationship to quality other than
what I have already quoted above.

```{r echo=FALSE}
ggplot(data = dataset, aes(density)) +
  geom_histogram(binwidth = .0005, col="black", fill="grey", alpha = .8) +
  scale_x_continuous(breaks=seq(0,1.1,.001))

summary(dataset$density)
```

Desity is almost perfectly bell-shaped in its distribution. Its value on the 
scale depends on the percent of alcohol and sugar content in the red wine. I 
have daubts that density can explain a lot of variance in quality but we will 
see how it performs in a correlation analysis.

```{r echo=FALSE}

ggplot(data = dataset, aes(pH)) +
  geom_histogram(binwidth = .05, col="black", fill="grey", alpha = .8) +
  scale_x_continuous(breaks=seq(2.5,4.5,.1))

summary(dataset$pH)
```

Very much like Desity, PH is relativly normal in its distribution. From the 
text file, which describes the variables and how the data was collected, I can 
gather that most wines are between 3-4 on the pH scale that spans from 0 (very 
acidic) to 14 (very basic). The data confirms this information and is 
illustrating that only lower PH values are suitable for wines. 

```{r echo=FALSE}

ggplot(data = dataset, aes(sulphates)) +
  geom_histogram(binwidth = .05, col="black", fill="grey", alpha = .8) +
  scale_x_continuous(breaks=seq(0,2,.1))

ggplot(data = dataset, aes(sulphates)) +
  geom_histogram(binwidth = .05, col="black", fill="grey", alpha = .8) +
  scale_x_continuous(breaks=seq(0,2,.1), limits = quantile(dataset$sulphates, c(0, .99)))

ggplot(data = dataset, aes(sulphates)) +
  geom_histogram(binwidth = .02, col="black", fill="grey", alpha = .8) +
  scale_x_log10(limits = quantile(dataset$sulphates, c(0, .99)))

summary(dataset$sulphates)
```

Sulphates are additives which can contribute to sulfur dioxide gas (S02) levels, 
wich acts as an antimicrobial and antioxidant. The first histogram shows a right
skewed distribution of how the data appears as a whole. Some outliers are 
stretching the distribution about 1/3 of where the majority of the data sits. 

I was attempting to transform the variable by limiting the histogram to 99% of 
the data. It appeared as if the distribution was still skewed to the right so 
that I decided to apply another transformation using a log10 method. 


```{r echo=FALSE}
ggplot(data = dataset, aes(alcohol)) +
  geom_histogram(binwidth = .25, col="black", fill="grey", alpha = .8) +
  scale_x_continuous(breaks=seq(8,15, .5))

ggplot(data = dataset, aes(alcohol)) +
  geom_histogram(binwidth = .25, col="black", fill="grey", alpha = .8) +
  scale_x_continuous(breaks=seq(8,15, .5), limits = quantile(dataset$alcohol, c(0, .99)))

summary(dataset$alcohol)

```

Alcohol, I would assume, is one of the most important variables in wine. There 
is a minimum of at least 8.5% of alcohol in every wine and a maximum of 15%. The 
distribution is right skewed peaking at 9.5% affecting the total mean to μ10.42. 

Unfortunately, I was not able to find an applicable transformation method that 
could bring the distribution to a more suitable shape for analysis. However, 
limiting the data to 99% excluding some of the outliers could fi some degree 
of the skeweness. However, the peak at 9.5% remains as the main explaination for 
its shape in the histogram.

# Univariate Analysis

### What is the structure of your dataset?

There are 1599 red wines included in the dataset with 12 features 
("fixed.acidity", "volatile.acidity", "citric.acid", "residual.sugar", 
"chlorides", "free.sulfur.dioxide", "total.sulfur.dioxide", "density", "pH", 
"sulphates", "alcohol", "quality". The variable quality is an integer whereas 
all other variables are numeric.

Other observations:

Higher number of wines from medium to higher quality. \
About 8% of red wines contain no citric acid. \
The median for residual.sugar is 2.200 and the max is 15.500.

### What is/are the main feature(s) of interest in your dataset?

The main feature in the dataset is quality. My intution leads me to assume that 
alcohol could be another main feature. However, I lack the necessary domain 
knowledge on chemcial properties and wine to identfy other main features from 
the analysis above and wihtout further statistical investigation. 

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

I have a feeling that the ratio of precense of one chemical property with 
the absence of another could possibly explain some of the variance in quality. 

### Did you create any new variables from existing variables in the dataset?

I lack the necessary domain knowledge on chemcial properties and wine to identfy 
appropriate opurtunities in the dataset that would make such operation useful. 

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

I have transformed quality to a factor as it is a categorical variable on a 
scale between 0 (very bad) and 10 (very excellent). Moreover, I have performed 
transformations on the following few variables in the dataset as many of them 
have been right skewed: citric.acid.sqr, residual.sugar.log10, chlorides.log10, 
free.sulfur.dioxide.log10, total.sulfur.dioxide.log10, sulphates.log10. The 
transformations have been sucessful exept for citric.acid.sqr. This variable 
still shows a bimodal wavepattern distribution with counts peaking around 0.25 
and 0.50. 

# Bivariate Plots Section

```{r echo=FALSE, Bivariate_Plots}
correlation_matrix = cor(dataset[,2:13])
round(correlation_matrix, 2)
#ggpairs(dataset)
pairs.panels(dataset[,2:14], smooth = TRUE, factor=5, stars=TRUE) 

# ggpairs(data=dataset[,2:13], title="GGPairs Matrix", 
#         lower = list(continuous = wrap("points", shape = I('.'))),
#         upper = list(combo = wrap("box", outlier.shape = I('.'))))
```

In order to get a better visual representation of the correlations, a heatmap of 
correlations is shown below. Warm colors indicate negative correlations whereas 
cold colors indicate positive correlations. 

```{r echo=FALSE}

corrplot(correlation_matrix, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)
```

At this point I was also wondering how the transformed variables would perform 
with each other and if there is any significant improvements. The pairs.panel 
below is intended to show just that.

```{r echo=FALSE}

#adding the transformed variables to the dataset:

dataset$citric.acid.sqr = sqrt(dataset$citric.acid)
dataset$residual.sugar.log10 = log10(dataset$residual.sugar)
dataset$chlorides.log10 = log10(dataset$chlorides)
dataset$free.sulfur.dioxide.log10 = log10(dataset$free.sulfur.dioxide)
dataset$total.sulfur.dioxide.log10 = log10(dataset$total.sulfur.dioxide)
dataset$sulphates.log10 = log10(dataset$sulphates)

# and re-run the correlation table and heatmap:

pairs.panels(dataset[,14:20], smooth = TRUE, factor=5, stars=TRUE) 

```

The heatmap above provides some intersting insides about the nature of the 
correlations without including any of the transformed variables. Moreover, the 
pairs.panels function from the psych package shows the relationship of the 
transformed variables. 

Generally, the bivariate correlations are rather weak and also the transformed 
variables are not necessarily performing better. The correlation coefficient 
improves slightly but not in a meaningful way that would justify further 
investigation.

The variables that correspond the strongest with quality are volatile.acidity 
and alcohol whereby the relationship itself is weak. The direction of 
volatile.acidity with quality is negative meaning that lower levels of 
volatile.acidity tend to have a better quality, which is shown in the boxplot 
below. 

```{r echo=FALSE}
ggplot(data = dataset, aes(quality.factor, volatile.acidity)) +
  geom_boxplot()

by(dataset$volatile.acidity, dataset$quality.factor,  summary)

ggplot(data = dataset, aes(volatile.acidity)) +
  geom_histogram(aes(fill = quality.factor), binwidth = .05, alpha = .8) +
  scale_x_continuous(breaks=seq(0,1.6,.1), limits = quantile(dataset$volatile.acidity, 
                                                             c(0, .99)))
```

On the other hand alcohol has a postive relationship with quality meaning, to 
put it cautiously, that its presence tend to influence a higher quality in wine 
as shown in the boxplot below. Interstingly, medium quality wine has the biggest
range in alcohol. I wonder if the combination with volatile.acidity can reveal
more insights?

```{r echo=FALSE}
ggplot(data = dataset, aes(quality.factor, alcohol)) +
  geom_boxplot()

by(dataset$alcohol, dataset$quality.factor, summary)

by(dataset$alcohol, dataset$quality.factor, range)

group_by(dataset, quality.factor) %>%
  summarise(
    count = n(),
    mean = mean(alcohol, na.rm = TRUE),
    sd = sd(alcohol, na.rm = TRUE)
  )

```

Looking at the relationships between the supporting variables I can see that 
total.sulfur.dioxide and free.sulfur.dioxide moderatly correlate with each 
other. According to the text file which describes the variables this makes a lot
of sense as free.sulfur.dioxide is part of the meassurement of 
total.sulfur.dioxide. I decided to not further investigate this relationship.

We can also see a weak to moderate negative correlation between density and 
alcahol. This makes sense as well as density of water is close to that of wine 
depending on the percent alcohol and sugar content. 

```{r echo=FALSE}

ggplot(data = dataset, aes(density, alcohol)) +
  geom_point()

ggplot(data = dataset, aes(density, fixed.acidity)) +
  geom_point()

```

The first scatterplot above shows the moderate negative relationship of density 
and alcohol of -0.50. That means as one variables increases, the other 
variable decreases. 

On the other hand density and fixed.acidity have a moderate to strong postive 
relationship of 0.67. That means fixed.acidity tends to increase with density.

```{r echo=FALSE}
ggplot(data = dataset, aes(fixed.acidity, pH)) +
  geom_point()

ggplot(data = dataset, aes(citric.acid, pH)) +
  geom_point()
```

Both scatterplots above attempt to explore the relationships between citric.acid 
(-0.54) and pH as well as fixed.acidity and pH (-0.68) found in the correlation 
matrix. The negative relationships can be seen in the dottet scatterplot-cloud. 
Again, I wonder how some of these variables related to acidity, pH, density and 
alcohol might support each other with regards to quality during the bivariate 
analysis.

I will focus on exploring the statistical relationship between alcohol and 
acidity on quality as a factor. I am showing an initial one-way ANOVA test below 
including plots checking homogeneity of variance and normality. I am going to 
start building a model with alcohol ~ factor(quality):

```{r echo=FALSE}

model1 = aov(alcohol ~ factor(quality), data = dataset)
summary(model1)
plot(model1, 1)
plot(model1, 2)
```

Points 653 and 145 are detected as outliers. I want to have a look at a similar 
model with volatile.acidity ~ factor(quality) before I further optimize the 
model.

```{r echo=FALSE}

model2 = aov(volatile.acidity ~ factor(quality), data = dataset)
summary(model2)
plot(model2, 1)
plot(model2, 2)
```

Points 128, 127 and 1300 are detected as outliers. I will remove them along with 
the outliers from alcohol ~ factor(quality) and re-run the calculation in order 
to minimize affects on normality and homogeneity of variance. 

```{r echo=FALSE}
rm_outliers_subset = subset(dataset, alcohol < quantile(alcohol, .99) & 
                              volatile.acidity < quantile(volatile.acidity, .99))

#rm_outliers_subset$quality = factor(quality)


model3 = aov(alcohol ~ quality, data = rm_outliers_subset)
summary(model3)
plot(model3, 1)
plot(model3, 2)
```

After subsetting to about 99% of the distribution in alcohol I could sucessfully 
remove all the relevant outliers.

```{r echo=FALSE}
model4 = aov(volatile.acidity ~ quality, data = rm_outliers_subset)
summary(model4)
plot(model4, 1)
plot(model4, 2)
```

After subsetting to about 99% of the distribution in volatile.acidity I could 
sucessfully remove all the relevant outliers.

```{r echo=FALSE}
aov_residuals.alcohol = residuals(object = model3 )
leveneTest(alcohol ~ factor(quality), data = rm_outliers_subset, center = mean)
shapiro.test(x = aov_residuals.alcohol)
```

After subsetting to about 99% of the distribution in alcohol and 
volatile.acidity I could sucessfully remove all the relevant outliers. Moreover, 
I ran a Levene's Test with a p-value that is not less than the significance 
level of 0.05. This means that there is no evidence to suggest that the variance 
across groups is statistically significantly different. Moreover, the 
Shapiro-Wilk test on the ANOVA residuals (W = 0.96, p = < 0.05) shows that, even 
after transforming the data, normality is violated.

```{r echo=FALSE}
aov_residuals.volatile.acidity = residuals(object = model4 )
leveneTest(volatile.acidity ~ factor(quality), data = rm_outliers_subset, center = mean)
shapiro.test(x = aov_residuals.volatile.acidity)
```

I also ran a Levene's Test with a p-value that is not less than the significance 
level of 0.05 for volatile.acidity.  Moreover, the Shapiro-Wilk test on the 
ANOVA residuals (W = 0.96, p = < 0.05) normality is violated indicating that
both independent variables population is not normal:

I am investigating further by using a Kruskal-Wallis rank sum test. This will 
allow me to see if levels of alcohol in lower level wines have identical data 
distributions than levels of alcohol in medium to higher quality wines without 
assuming the data to have normal distribution at a .05 significance level.

```{r echo=FALSE}
kruskal.test(alcohol ~ quality, data = rm_outliers_subset)
pairwise.wilcox.test(rm_outliers_subset$alcohol, rm_outliers_subset$quality,
                 p.adjust.method = "BH")
```

After seeing that the test is signifaicant I decided to apply the same method 
for volatile.acidity.

```{r echo=FALSE}
kruskal.test(volatile.acidity ~ quality, data = rm_outliers_subset)
pairwise.wilcox.test(rm_outliers_subset$volatile.acidity, rm_outliers_subset$quality,
                 p.adjust.method = "BH")

```

Interestingly, this data reveals that pretty much only the medium quality levels 
differ from each other. We see significant differences from 5-4, from 5-6 and 
from 6-7. Moreover, the variable from the very top correlate with the bottom 
quality categories. 


# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

Overall, the relationships between the feature of interest and the other 
features in the dataset present themself rather weak. Even the transformation of
some of the variables could not really make a difference to change this 
perception. The strongest relationship with quality has alcohol meaning that 
higher levels of alcohol moderately correlate with higher quality ratings. 

The variable volatile.acidity has the second strongest relationship with 
quality and among all the other variables in the dataset it is the only one left 
worth mentioning in the context of quality.  

http://www.sthda.com/english/wiki/correlation-matrix-a-quick-start-guide-to-analyze-format-and-visualize-a-correlation-matrix-using-r-software

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?

Alcohol and volatile.acidity have the strongest relationshop with quality. With 
significant p-values from both performed Kruskal-Wallis rank sum tests I 
can assume that the variation of alcohol and volatile.acidity among different 
quality categories is much larger than the variation of alcohol and 
volatile.acidity within each quality category. Hence I could conclude that there 
is a significant relationship between quality categories and 
alcohol as well as volatile.acidity.

# Multivariate Plots Section

My first intention for the multivariate section was to color some of the 
scatterplots from above to see if quality will reveal some additional patterns. 

```{r echo=FALSE, Multivariate_Plots}

ggplot(data = dataset, aes(density, alcohol)) +
  geom_point(aes(color = quality.factor)) +
  geom_smooth(se = FALSE, method = lm) +
  ylab("alcohol (% by volume)") +
  xlab("density (g / cm^3)") +
  labs(color="quality score (0-10)") 

ggplot(data = dataset, aes(density, fixed.acidity)) +
  geom_point(aes(color = quality.factor)) +
  geom_smooth(se = FALSE, method = lm) +
  ylab("fixed acidity (tartaric acid - g / dm^3)") +
  xlab("density (g / cm^3)") +
  labs(color="quality score (0-10)")

ggplot(data = dataset, aes(density, volatile.acidity)) +
  geom_point(aes(color = quality.factor)) +
  geom_smooth(se = FALSE, method = lm) +
  ylab("volatile acidity (acetic acid - g / dm^3)") +
  xlab("density (g / cm^3)") +
  labs(color="quality score (0-10)")


```

The coloring with quality.factor.

```{r echo=FALSE}
ggplot(data=dataset, aes(x=density, y=alcohol, col=quality.factor, 
                         size=fixed.acidity)) + 
  geom_point(alpha=0.25) +
  ylab("alcohol (% by volume)") +
  xlab("density (g / cm^3)") +
  labs(color="quality score (0-10)") +
  labs(size="fixed acidity (g / dm^3)") 

ggplot(data=dataset, aes(x=density, y=alcohol, col=quality.factor, 
                         size=citric.acid)) + 
  geom_point(alpha=0.5) +
  ylab("alcohol (% by volume)") +
  xlab("density (g / cm^3)") +
  labs(color="quality score (0-10)") +
  labs(size="citric acid (g / dm^3)") 

ggplot(data=dataset, aes(x=density, y=alcohol, col=quality.factor, 
                         size=sulphates)) +
         geom_point(alpha=0.5) +
  ylab("alcohol (% by volume)") +
  xlab("density (g / cm^3)") +
  labs(color="quality score (0-10)") +
  labs(size="sulphates (g / dm3)") 

```

Moreover, I was trying to explore additional variables with density, alcohol and
quality.

```{r echo=FALSE}
ggplot(data = dataset, aes(fixed.acidity, pH)) +
  geom_point(aes(color = quality.factor)) +
  ylab("pH (1-14)") +
  xlab("fixed acidity (g / dm^3)") +
  labs(color="quality score (0-10)")

ggplot(data = dataset, aes(citric.acid, pH)) +
  geom_point(aes(color = quality.factor)) +
  ylab("pH (1-14)") +
  xlab("citric acid (g / dm^3)") +
  labs(color="quality score (0-10)")

```

I have also highlighted quality.factor in the colors with fixed.acidity and pH
as well as citric.acid and pH.

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

My idea for this part of the assigment was to add the quality factor to the 
existing plots I created. The coloring with quality.factor indicates a tendency 
that higher qualities tend to have more alcohol and less density whereas higher 
desity and lower levels of alcohol indicate more medium to lower quality wines.

It also seems like as if quality is slightly layered with fixed.acidity and 
density. This could mean that higher quality wines tend to have higher amounts 
of fixed.acidity while having slighty lower amounts of density as medium quality
wines. 

Moreover, I was trying to explore additional variables with density, alcohol and
quality that I had originally found being minimally connected with quality from 
the cor table. However, the plots do not really reveal any new secrets worth 
following up on. I have also highlighted quality.factor in the colors with 
fixed.acidity and pH as well as citric.acid and pH. This, however, illustrates 
how neither of these relationships is tied to quality in any meaningful and 
visual way. 

### Were there any interesting or surprising interactions between features?

Unfortunately, I haven't been able to un-cover any new insights in this section
that really suprised me. I can imagine that a more sophisticated familiarity 
with the chemical properties in the dataset could result in better insights.

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

I have not created any models beyond the bivariate models through the lack of 
knowledge of how to choose an appropriate statistical procedure. Further 
investigation would be needed.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}
ggplot(rm_outliers_subset, aes(factor(quality), alcohol)) + 
  geom_boxplot(varwidth=T, aes(fill=quality)) + 
  geom_smooth(method = "loess", se=FALSE, color="plum", aes(group=1)) +
    labs(title="Box plot", 
         subtitle="Alcohol grouped by Quality of Wine",
         x="Quality Score (0-10)",
         y="Alcohol (% by volume)") +
  labs(fill ="Quality Score (0-10)")
```

### Description One

I choose a box plot for my first two plots as I believe it is the best way to
visualize the most important data in this assignment. The box plot show the 
distributions of alcohol within different quality groups, along with the median, 
range and outliers. The width of the boxes is proportional to the number of 
observation it contains. We can see that most of the wines are distributed 
between 5 and 6. These two categories also have the greatest variance, which is 
especially effecting the category 5. Overall, however, we can see a clear 
trendline underpinning the staticial results that there is a significant 
difference in the distribution of alcohol in the among some of the categories.

### Plot Two
```{r echo=FALSE, Plot_Two}
ggplot(rm_outliers_subset, aes(factor(quality), volatile.acidity)) + 
  geom_boxplot(varwidth=T, aes(fill=quality)) + 
  geom_smooth(method = "loess", se=FALSE, color="plum", aes(group=1)) +
    labs(title="Box plot", 
         subtitle="Volatile Acidity grouped by Quality of Wine",
         x="Quality Score (0-10)",
         y="Volatile Acidity (g / dm^3)")  +
  labs(fill ="Quality Score (0-10)")

```

### Description Two

I also choose a box plot for my second plot in order to highlight the second most
important variable explaining some of the variance in quality. Just like the 
first box it shows the distributions of the variable within different quality 
groups, along with the median, range and outliers. We can see that most of the 
wines are distributed between 5 and 6 and that these two categories have the 
greatest variance. Overall the trendline is more clear than in the first plot as
these outliers are a little bit more distributed. The plot is also underpinning 
the staticial results from the Kruskal-Wallis test showing that there is a 
significant difference in the distribution of volatile.acidity among some of the 
categories.


### Plot Three
```{r echo=FALSE, Plot_Three}

ggplot(data=rm_outliers_subset, aes(x=density, y=alcohol, col=quality, 
                         size=volatile.acidity)) + 
  geom_point(alpha=0.25) +
  scale_x_continuous(breaks=seq(0,1.1,.001)) +
      labs(title="Scatterplot", 
         subtitle="Alcohol with Density, colored by quality and sized by volatile acidity",
         x="Density (g / cm^3)",
         y="Volatile Acidity (g / dm^3)")  +
  labs(color ="Quality Score (0-10)",
       size="Volatile Acidity (g / dm^3)")
```

### Description Three

The third and last plot is a summarisation of how I would want to further 
explore the model explaining the variance in quality. It is an attempt to 
combine the most important variable in the exploratiry data analysis. Alcohol and 
quality have the strongest relationship in the analysis so far. My thought was 
to use this relationship and expand with density as density correlate very well
back with alcohol. The less dense the wine is the more alcohol it seems to 
contain as the negative relationship suggests. I was coloring this relationship
with the factor quality to see if this can further clarify. It looks like as if
lower quality wines tend to have less alcohol, which in turn results in higher 
density. The size of the points in the scatterplot is determined by the level of
volatile.acidity as I was hoping it would add to the overall significance of the
plot. Unfortunately, I don't see the size by volatile.acidity to be that 
informative.
------

References:

R for Data Science Book by Garrett Grolemund and Hadley Wickham \ 
https://www.r-bloggers.com/ \
http://www.sthda.com/english/wiki/one-way-anova-test-in-r
